---
title: "R Notebook_3"
output: html_document
html_notebook: default
df_print: paged
date: August 4, 2019
---

```{r setup, include = F}
chooseCRANmirror(graphics = F, ind = 1)
knitr::opts_chunk$set(echo = T)
```

# Chapter 3 Probability

## Random Sampling in R

* sample() function

```{r}
# the function randomly sorts the values by default
sample(c(1:10))
```

* Sampling with replacement

```{r}
# "replace" argument
sample(c(1:10), replace = T)
```

* Sampling a fixed number of values

```{r}
# "size" argument
sample(c(1:10), size = 5, replace = T)
```

* Setting the seed of the random number generator

```{r}
set.seed(12345)
sample(c(1:10), size = 5, replace = T)
```

* Simulating a random process

```{r}
# rolling a fair die; estimate the probability of getting a "1"
size <-  100
x <- sample(c(1:6), size = size, replace = T)
mean(x == 1)
```

* Adjusting probabilities

```{r}
# defective rate is 2%
size <-  100
x <- sample(c('G', 'D'), size = size, replace = T, prob = c(0.98, 0.02))
mean(x == 'D')
```

* Randomly selecting rows in a dataset

```{r}
# defective rate is 2%
size <-  5
D <-  datasets::cars
subD <-  D[sample(nrow(D), size = size),]
subD
```

## Generating Random Numbers

* Simulating a uniform distribution

```{r}
size <-  5
# random numbers between 0 and 1
x <- runif(size)
x
```


```{r}
size <-  5
# random numbers between a and b
x <- runif(size, 10, 20)
x
```

* Simulating a normal distribution

```{r}
size <-  5
x <- rnorm(size, mean = 10, sd = 2)
x
```

* Simulating a binomial distribution

```{r}
size <-  5
# n = 150, p = 0.05, the number of successes is returned
x <- rbinom(size, 150, 0.05)
x
```

