---
title: "R Notebook_1"
output:
  html_document:
    df_print: paged
---

# Chapter 1 Introduction to Data

## R and RStudio

* R is a programming language.
* RStudio is a R programming environment (IDE). 
  + Panels: Console, Source, Environment, Plots
  + Source: Run (*Ctrl+Enter*); Commenting
  + Console: Automatic completion (Tab); Retrieving previous commands; Clear the console (*Ctrl+L*); Working directory
* A R Notebook is an R markdown document with R code chunks that can be executed with output visible immediately beneath the input. 
* To execute a code chunk, place your cursor inside it and press *Ctrl+Shift+Enter*. 

```{r}
plot(cars)
```

## Package `openintro`

* A package bundles together code, data and documentation.
* A repository is where packages are stored. 
* CRAN - Comprehensive R Archive Network
* [openintro](https://github.com/OpenIntroStat/openintro-r-package)

```{r}
# download and install RTools 
# https://cran.r-project.org/bin/windows/Rtools/
# install.packages("glue")
install.packages("devtools")
library(devtools)
install_github("openintrostat/openintro-r-package")
```

* The following datasets are included in the package. 

```{r}
data(package = "openintro")
```

## Read and Analyze Datasets

* Load the package

```{r}
library(openintro)
```

* Understanding a dataset

```{r}
help("gpa")
```

```{r}
# str() function displays the structure
str(gpa)
```

```{r}
# dim() displays the dimension
dim(gpa)
nrow(gpa)
ncol(gpa)
```

```{r}
# names() returns the variable names
names(gpa)
```

```{r}
# a factor variable
class(gpa$gender)
levels(gpa$gender)
```

* Accessing data

```{r}
# the first six observations
head(gpa)
```

```{r}
# the last six observations
tail(gpa)
```

```{r}
# indices start with 1
gpa[1, 1]
```

```{r}
# first row
gpa[1,]
```

```{r}
# second column
gpa[, 2]
```

* Subsetting data

```{r}
# first and third rows/first, third and fifth columns
gpa[c(1, 3), -c(2, 4)]
```

```{r}
# variable gpa
gpa$gpa
```

```{r}
# 7 observations
gpa[4:10,]
```

```{r}
# a variable
gpa$out[1:5]
```

```{r}
# variable names
gpa[c(1,3,5), c("gpa", "out")]
```

```{r}
# a comparison expression
gpa$gpa > 3.5
```

```{r}
# add a filter
gpa$out[gpa$gpa > 3.5]
```

```{r}
# subset() returns a subset
subset(gpa, gpa > mean(gpa) & out > mean(out))
```

## Scatter Plots

```{r}
# gpa and studyweek
plot(gpa$studyweek, gpa$gpa, main = "Study Hours vs. GPA", xlab = "Weekly Study Hours", ylab = "GPA", xlim= c(0, max(gpa$studyweek)), ylim = c(2, 5), pch = 20, col = COL[1])
```

```{r}
# median income and homeownership rate
plot(county$multi_unit, county$homeownership, main = "Multi_Unit vs. Homeownership", xlab = "Perent of Units in Multi_unit Structures", ylab = "Homeownership Rate", xlim = c(0, max(county$multi_unit)), ylim = c(0, 100), pch = 20, col = COL[1])
```
