---
title: "R Lab 1"
output: html_document
html_notebook: default
df_print: paged
date: August 1, 2019
---

```{r setup, include = F}
chooseCRANmirror(graphics = F, ind = 1)
knitr::opts_chunk$set(echo = T)
```

# Lab 1 Introduction to R

To get you started, run the following R code chuck. 

```{r load-abrbuthnot-data, eval=TRUE}
source("http://www.openintro.org/stat/data/arbuthnot.R")
```

This command instructs R to access the OpenIntro website and fetch some data: the Arbuthnot baptism counts for boys and girls. A description of the dataset is available at the [OpenIntro website](https://www.openintro.org/stat/data/?data=arbuthnot). 

## The Data: Dr. Arbuthnot's Baptism Records

R has stored Arbuthnot's data in a *data frame*. The name of the data frame is `arbuthnot`. 

```{r}
####################################
# TODO 1:                          #   
#                                  #
# Write R code in this code chunk  #
# to output the number of rows and #
# the number of columns in the     #
# arbuthnot dataset.               #
#                                  #
####################################

```

```{r}
####################################
# TODO 2:                          #   
#                                  #
# Write R code in this code chunk  #
# to output the names of all       #
# variables.                       #
#                                  #
####################################

```

```{r}
####################################
# TODO 3:                          #   
#                                  #
# Write R code in this code chunk  #
# to output the number of boys     #
# baptized in the first 3 years.   #                       #
#                                  #
####################################

```

```{r}
####################################
# TODO 4:                          #   
#                                  #
# Write R code in this code chunk  #
# to output the numbers of boys    #
# and girls baptized, respectively,#
# in the first 3 years.            #
#                                  #
####################################

```

```{r}
####################################
# TODO 5:                          #   
#                                  #
# Write R code in this code chunk  #
# to output the years in which     #
# boys baptized outnumbered the    #
# yearly average of boys.          #
#                                  #
####################################

```

```{r}
####################################
# TODO 6:                          #   
#                                  #
# Write R code in this code chunk  #
# to create a scatter plot for     #
# the number of boys baptized vs.  #
# the number of girls baptized.    #
#                                  #
####################################

```


```{r}
####################################
# TODO 7:                          #   
#                                  #
# Write R code in this code chunk  #
# to create a plot that shows how  #
# the number of girls              #
# evolved over years. To connect   #
# the points with lines, you may   #
# add "l" as the third argument    #
# to the plot() function.          #
#                                  #
####################################

```


```{r}
####################################
# TODO 8:                          #   
#                                  #
# Write R code in this code chunk  #
# to compute the total number of   #
# baptisms each year.              #
#                                  #
####################################

```

```{r}
####################################
# TODO 9                           #  
#                                  #
# Write R code in this code chunk  #
# to create a plot that shows how  #
# the proportion of boys evolved   #
# over years.                      #
#                                  #
####################################

```