---
title: "R Notebook_4"
output: html_document
html_notebook: default
df_print: paged
date: August 8, 2019
---

```{r setup, include = F}
chooseCRANmirror(graphics = F, ind = 1)
knitr::opts_chunk$set(echo = T)
```

# Chapter 4 Distributions of Random Variables

## Simulating the Binomial Distribution

The binomial distribution is used to describe the number of successes in a fixed number of Bernoulli trials (the outcome of each trial is either a success or a failure). 

* Sampling from a large population

```{r}
pop.size = 1E4
# probability of a success
prob.y = 0.4
sp.size = 10
# number of samples
sps <- 1E3
# construct the population
pop.entries <- c(rep("y", pop.size * prob.y), rep("n", pop.size * (1 - prob.y)))
# computing the number of successes in each sample
set.seed(12345)
# generate 1000 random samples with a sample size of 10
sp.ones <-  replicate(sps, sum(sample(pop.entries, sp.size, replace = F) == "y"))
# analyzing the simulation results
hist(sp.ones)
prop.table(table(sp.ones))
mean(sp.ones)
sd(sp.ones)
```


* Sampling from a small population

```{r}
pop.size <-  50
# construct the population
pop.entries <- c(rep("y", pop.size * prob.y), rep("n", pop.size * (1 - prob.y)))
# sampling without replacement
set.seed(12345)
sp.ones <- replicate(sps, sum(sample(pop.entries, sp.size, replace = F) == "y"))
# analyzing the simulation resultshist(sp.ones)
prop.table(table(sp.ones))
mean(sp.ones)
sd(sp.ones)
```

## Normal Distribution

Suppose random variable X is normally distributed with mean 50 and sd 12.

* R functions

```{r}
x <- 65
mean <- 50
sd <- 12
# density function value f(x)
dnorm(x, mean, sd)
# P(X < x)
pnorm(x, mean, sd)
# P (X > x)
pnorm(x, mean, sd, lower.tail = F)
# P(0.5x < X < x)
pnorm(x, mean, sd) - pnorm(0.5*x, mean, sd)
# y such that P(X < y) = 0.75
prob <- 0.75
qnorm(prob, mean, sd)
# y such that P(X > y) = 0.9
prob <- 0.9
qnorm(prob, mean, sd, lower.tail = F)
# generate ten random numbers
size <- 10
rnorm(size, mean, sd)
```

* Evaluating the normal distribution

```{r}
size  <-  20
normdata <- rnorm(size, mean, sd)
# producing a normal qqplot (quantile-quantile plot) to check # normality; the distribution is normal if the points are close to form a straight line.  
qqnorm(normdata)
# add a "theoretical" line
qqline(normdata)
```