---
title: "R Notebook_7"
output: html_document
html_notebook: default
df_print: paged
date: August 21, 2019
---

```{r setup, include = F}
chooseCRANmirror(graphics = F, ind = 1)
knitr::opts_chunk$set(echo = T)
```

# Chapter 7 Inference for Numerical Data

## A Population Mean

Chapter 7 Exercise 1

n = 50
xbar = 7.0
s = 1.4
alpha = 0.05

* One-sided Hypothesis Testing 
null: mu = 6.5
alternative: mu > 6.5

```{r}
# sample size
n <- 50
# sample mean & standard deviation
xbar <- 7.0
s <- 1.4
# significance level
alpha <- 0.05

# hypothesized parameter
muo <- 6.5
# t-statistic
t <- (xbar - muo) / (s / sqrt(n))
# p-value P(T > t)
pv <- pt(t, n - 1, lower.tail = F)
pv
```

* 95% Confidence Interval 

```{r}
# sample size
n <- 50
# sample mean & standard deviation
xbar <- 7.0
s <- 1.4
# significance level
alpha <- 0.05

# standard error
se <- s / sqrt(n)
# margin of error
me <- qt(alpha / 2, n - 1, lower.tail = F) * se
# lower confidence limit
lcl <- xbar - me
# upper confidence limit
ucl <- xbar + me
lcl
ucl
```

## Difference of Two Population Means

Chapter 7 Exercise 2

men_x = {8, 7, 10, 5, 7}
women_x = {9, 5, 12, 8}
alpha = 0.05

* One-sided Hypothesis Testing 

Hypotheses
null: mu1 - mu2 = 0
alternative: mu1 - mu2 < 0

```{r}
# data sets
men_x <- c(8, 7, 10, 5, 7)
women_x <- c(9, 5, 12, 8)

# sample sizes
n1 <- length(men_x)
n2 <- length(women_x)

# sample means
xbar1 <- mean(men_x)
xbar2 <- mean(women_x)
s1 <- sd(men_x)
s2 <- sd(women_x)

# sample difference
xbardiff <- xbar1 - xbar2
# standard error
se <- sqrt(s1 ** 2 / n1 + s2 ** 2 / n2)
# test statistic
t <- xbardiff / se

# margin of error
df <- min(n1 - 1, n2 - 1)
# pv = P(T < t)
pv <- pt(t, df)
pv
```


* 95% Confidence Interval 

```{r}
# data sets
men_x <- c(8, 7, 10, 5, 7)
women_x <- c(9, 5, 12, 8)

# sample sizes
n1 <- length(men_x)
n2 <- length(women_x)

# sample means
xbar1 <- mean(men_x)
xbar2 <- mean(women_x)
s1 <- sd(men_x)
s2 <- sd(women_x)

# significance level
alpha <- 0.05

# sample difference
xbardiff <- xbar1 - xbar2
# standard error
se <- sqrt(s1 ** 2 / n1 + s2 ** 2 / n2)

# margin of error
df <- min(n1 - 1, n2 - 1)
me <- qt(alpha / 2, df, lower.tail = F) * se

# lower confidence limit
lcl <- xbardiff - me
# upper confidence limit
ucl <- xbardiff + me
lcl
ucl
```

## Analysis of Variance

Chapter 7 Exercise 3

3 groups

Sample size  160   205   64
Sample mean  0.32  0.318  0.302
Smaple SD    0.043 0.038  0.038

```{r}

# number of groups
k <- 3

# sample sizes
szs <- c(160, 205, 64)
n <- sum(szs)

# sample means
means <- c(0.32, 0.318, 0.302)
xbar <- sum(szs * means) / n  

# sample sds
sds <- c(0.043, 0.038, 0.038)

# sum of squares between groups
SSG <- sum(szs * (means - xbar) ** 2)

# mean square between groups
MSG <- SSG / (k - 1)

# sum of squared errors
SSE <- sum((szs - 1) * sds ** 2)

# mean squared error
MSE <- SSE / (n - k)

# test statistic 
f <- MSG / MSE
dfg <- k - 1
dfe <- n - k

# pv = P(F > f)
pv <- pf(f, dfg, dfe, lower.tail = F)
pv
```