---
title: "R Notebook_2"
output: html_document
html_notebook: default
df_print: paged
date: July 31, 2019
---

```{r setup, include = F}
chooseCRANmirror(graphics = F, ind = 1)
knitr::opts_chunk$set(echo = T)
```

# Chapter 2 Summarizing Data

* Load the openintro Package

```{r}
library(openintro)
```

## Visualizing Numerical Data

* Dot plots

```{r}
x <- loan50$interest_rate
dotPlot(x, xlim = c(0.9 * min(x), 1.05 * max(x)), ylim = c(0.96, 1.05), xlab = "Interest rate", ylab = "", at = 1.007, pch = 20, col = COL[1], cex = 2, axes = F)
abline(h = 0.98)
AxisInPercent(1, pretty(c(0, x)))
Avg = mean(x)
polygon(Avg + c(-1, 0, 1) * 1, c(0.96, 0.98, 0.96), col = COL[4], border = COL[4])
```

* Stacked Dot plots

```{r}
x <- loan50$interest_rate
# round x to the nearest percent
y <- round(x)
stripchart(y, method = "stack", offset = 1/3, xlim = c(0.9 * min(y), 1.05 * max(y)), ylim = c(0, 10), xlab = "Interest rate, Rounded to the Nearest Percent", ylab = "", at = 1, pch = 20, col = COL[1], cex = 2, axes = F)
abline(h = 0.2)
AxisInPercent(1, pretty(c(0, y)))
Avg = mean(x)
polygon(Avg + c(-2, 0, 2) * 1, c(-1, 0.2, -1), col = COL[4], border = COL[4])
```

* Histograms

```{r}
hist(x, breaks = 5, xlab = "Interest Rate", ylab = "Frequency", main = "Counts for the binned interest rate data", xlim = c(0, 1.25 * max(x)), ylim = c(0, 30), col = COL[1], border = COL[5], axes = F)
abline(h = 0)
axis(2)
AxisInPercent(1, pretty(x))
```

* Box plots

```{r}
boxplot(x, ylab = "Interest Rate", xlim = c(0.3, 3), axes = F, ylim = range(x) + sd(x) * c(-1, 1) * 0.5)
AxisInPercent(2, c(0, pretty(x)))
arrows(2, min(x) + 1, 1.35, min(x), length = 0.08)
text(2, min(x) + 1, 'lower whisker', pos = 4)

arrows(2, quantile(x, 0.25) + sd(x) / 7,
       1.35, quantile(x, 0.25),
       length = 0.08)
text(2, quantile(x, 0.25) + sd(x)/6.5,
     expression(Q[1]~~'(first quartile)'), pos = 4)

m <- median(x)
arrows(2, m + sd(x) / 5, 1.35, m, length = 0.08)
text(2,m + sd(x) / 4.7, 'median', pos = 4)

q <- quantile(x, 0.75)
arrows(2, q + sd(x) / 4, 1.35, q, length = 0.08)
text(2, q + sd(x) / 3.8,
     expression(Q[3]~~'(third quartile)'), pos = 4)

arrows(2, rev(sort(x))[3] - sd(x) / 4,
       1.35, rev(sort(x))[3], length = 0.08)
text(2, rev(sort(x))[3] - sd(x) / 3.8,
     'upper whisker', pos = 4)

z <- quantile(x, 0.75) + 1.5 * IQR(x)
arrows(2, z - 0.1 * sd(x),
       1.35, z, length = 0.08)
lines(c(0.72, 1.28), rep(z, 2),
      lty = 3, col = '#00000066')
text(2, z - 0.1 * sd(x),
     'max whisker reach', pos = 4)

m <- rev(tail(sort(x), 5))
s <- m[1] - 0.3 * sd(m)
arrows(2, s, 1.1, m[1] - 0.2, length = 0.08)
arrows(2, s, 1.1, m[2] + 0.3, length = 0.08)
text(2, s, 'suspected outliers', pos = 4)

```

* Transforming data

```{r}
# original population data 
pop = county$pop2017
hist(pop, breaks = 12, main = "", xlab = "Population (M = million)", ylab = "Frequency", axes = F, col = COL[1], border = COL[5])
axis(2)
axis(1, at = pretty(pop), paste0(pretty(pop/1.0e6), "M"))
```


```{r}
# log-transformed population data 
hist(log(pop, 10), breaks = 12, main = "", xlab = expression(log[10]*"(Population)"), ylab = "Frequency", axes = F, col = COL[1], border = COL[5])
axis(2)
axis(1)
```

## Summary Measures for Numerical Data

* summary() function

```{r}
summary(x)
```

* Measures of central location

```{r}
# mean(x, na.rm = T) 
mean(x)
# median(x, na.rm = T) 
median(x)
# mode
Mode <- function(d) {
  ud <- unique(d)
  ud[which.max(tabulate(match(d, ud)))]
}
Mode(x)
```

* Measures of variability

```{r}
# range
range(x)
min(x)
max(x)

# sample var
var(x)
# population var
var(x) * (length(x) - 1) / length(x)

# sample std
sd(x)
# population std
sd(x) * sqrt((length(x) - 1) / length(x))
```

* Measures of relative standing

```{r}
# quartiles
quantile(x)

# IQR
IQR(x)

# Percentile
quantile(x, 0.2)
```

## Visualizing Categorical Data

* A One-way Contingency Table

```{r}
ho <-  loans_full_schema$homeownership
ho <-  droplevels(ho)
tbl1 <-  addmargins(table(ho), margin = 1)
names(tbl1) <-  tolower(names(tbl1))
names(tbl1)[4] <- "total"
tbl1 <-  tbl1[c("rent", "mortgage", "own", "total")]
tbl1
```

* A Two-way Contingency Table

```{r}
app <-  loans_full_schema$application_type
app <-  droplevels(app)
tbl2 <-  table(app, ho)
colnames(tbl2) <-  tolower(colnames(tbl2))
tbl2 <- tbl2[, c("rent", "mortgage", "own")]
tbl2
tbl3 <- addmargins(tbl2, margin = c(1, 2))
colnames(tbl3)[4] <- "total"
rownames(tbl3)[3] <- "total"
tbl3
```

* A two-way contingency table of proportions

```{r}
tbl2.prop = prop.table(tbl2)
tbl3.prop = addmargins(tbl2.prop, margin = c(1,2))
colnames(tbl3.prop)[4] <- "total"
rownames(tbl3.prop)[3] <- "total"
round(tbl3.prop, 3)
```

* A two-way contingency table of row proportions

```{r}
tbl2.prop = prop.table(tbl2, margin = 1)
tbl3.prop = addmargins(tbl2.prop, margin = 2)
colnames(tbl3.prop)[4] <- "total"
round(tbl3.prop, 3)
```

* A two-way contingency table of column proportions

```{r}
tbl2.prop = prop.table(tbl2, margin = 2)
tbl3.prop = addmargins(tbl2.prop, margin = 1)
rownames(tbl3.prop)[3] <- "total"
round(tbl3.prop, 3)
```

* Bar plots

```{r}
barplot(tbl1[1:3],
        axes = TRUE,
        xlab = '',
        ylab = 'Frequency',
        main = '',
        ylim = c(0, 1.2 * max(tbl1[1:3])),
        col = COL[1])
abline(h = 0)
mtext("Homeownership", 1, 2.4)
```

* Stacked bar plots

```{r}
ylim <- c(0, 1.2 * max(apply(tbl2, 2, sum)))
barplot(apply(tbl2, 2, sum),
    col = COL[3],
    ylim = ylim)
tblTemp <- tbl2[1,]
names(tblTemp) <- NULL
barplot(tblTemp,
    col = COL[1],
    add = TRUE,
    axes = FALSE)
abline(h = 0)
legend("topright",
    fill = COL[c(3,1)],
    legend = c("joint", "individual"))
par(las = 0)
mtext("Frequency", 2, 2.9)
```

* Side-by-side bar plots

```{r}
barplot(tbl2,
    ylim = c(0, max(tbl2)),
    col = COL[c(1, 3)], beside = TRUE)
abline(h = 0)
legend("topright",
    fill = COL[c(3,1)],
    legend = c("joint", "individual"))
par(las = 0)
mtext("Frequency", 2, 2.9)
```

* Standardized bar plots

```{r}
barplot(apply(tbl2, 2, sum) / apply(tbl2, 2, sum), col = COL[3])
tbl2.prop = prop.table(tbl2, margin = 2)
tblTemp <- tbl2.prop[1, ]
names(tblTemp) <- NULL
barplot(tblTemp,
    col = COL[1],
    add = TRUE,
    axes = FALSE)
legend(2.65, 0.3,
    fill = COL[c(3,1)],
    legend = c("joint", "individual"),
    bg = "white")
abline(h = 0)
par(las = 0)
mtext("Proportion", 2, 2.9)
```

* Mosaic plots

```{r}
# a one-variable mosaic plot
mosaicplot(tbl1[1:3],
           main = '',
           xlab = '',
           ylab = '',
           off = 4,
           col = COL[c(2, 1, 4)])

# a two-variable mosaic plot
mosaicplot(t(tbl2), xlab = "", ylab = "", main = "", col = COL[c(1, 2, 4)])
```

* Pie chars

```{r}
pie(table(ho), col = COL[c(1, 2, 4)])
```

* Side-by-side box plots

```{r}
boxplot(loans_full_schema$loan_amount ~ droplevels(loans_full_schema$homeownership), col = COL[c(1, 2, 4)], ylab = "Loan Amount", xlab = "Homeownership")
```

* Hollow histograms

```{r}
cc  <- county
popch <- sign(cc$pop2017 - cc$pop2010 - 0.5)
# sum(is.na(popch))
pov <- cc$median_hh_income
popch[popch == 1] <- "Gain"
popch[popch == -1] <- "No Gain"
xlim <- range(pov[popch == 'No Gain'], na.rm = TRUE)
histPlot(pov[popch == 'No Gain'],
         breaks = 50,
         col = '#ffffff00',
         border = COL[4],
         probability = TRUE,
         xlim = xlim,
         xlab = 'Median Household Income',
         ylab = '',
         hollow = TRUE,
         axes = FALSE,
         lty = 3,
         lwd = 4)
histPlot(pov[popch == 'No Gain'],
         breaks = 50,
         col = '#ffffff00',
         border = COL[4],
         probability = TRUE,
         add = TRUE,
         hollow = TRUE,
         lty = 3,
         lwd = 2)
histPlot(pov[popch == 'No Gain'],
         breaks = 50,
         col = '#ffffff00',
         border = COL[4],
         probability = TRUE,
         add = TRUE,
         hollow = TRUE,
         lty = 3,
         lwd = 1)
histPlot(pov[popch == 'Gain'],
         breaks = 50,
         col = '#ffffff00',
         border = COL[1],
         probability = TRUE,
         add = TRUE,
         hollow = TRUE,
         lty = 1,
         lwd = 2)
AxisInDollars(1, at = pretty(xlim))
legend('topright',
       col = COL[c(1,4)],
       lty = c(1,3),
       lwd = c(2,2.8),
       legend = c('Gain', 'No Gain'))
legend('topright',
       col = c(rgb(0,0,0,0), COL[4]),
       lty = c(1, 3),
       lwd = c(2,1.4),
       legend = c('Gain', 'No Gain'),
       bg = rgb(0,0,0,0),
       box.col = rgb(0,0,0,0),
       text.col = rgb(0,0,0,0))
legend('topright',
       col = c(rgb(0,0,0,0), COL[4]),
       lty = c(1, 3),
       lwd = c(2,0.7),
       legend = c('Gain', 'No Gain'),
       bg = rgb(0,0,0,0),
       box.col = rgb(0,0,0,0),
       text.col = rgb(0,0,0,0))
```